
---
title: "NFL Quarterback Statistics"
author: "Michael Saenz"
date: "05/02/2021"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Key Quarterback Statistics

Over the past few years, statistics and data have played an increasingly larger role in sports whether it is teams looking to get an advantage by selecting the best player or by regular people looking for help in picking players for fantasy leagues.  

People are using data points and statistics to help them out and make better informed decisions that are based on supporting rationale.  For football players, in particular Quarterbacks, there are a few key data points that are collected each year on every player that when analyzed, becomes easier to identify players that outperform their peers. These data points / stats that we take a closer look at below include total yards per season, completion percentage, total touchdowns per season, the passer ratings, and the quarterback rating.  
These stats can also be found at [ESPN.com](https://www.espn.com/nfl/stats/player/_/table/passing/sort/passingYards/dir/desc). 

 

#### **Quarterback Stats Over the Years: Visualization**  
When you look at quarterbacks in the NFL and are trying to figure out who the next top player might be or how they match up to the greats, it helps to focus on the key stats over the last decade.  Below is an interactive graph that helps you visualize the trends in a few of the key quarterback stats over the past 13 years.  

The Data points / Stats utilized contain the average of the top 10 quarterbacks from each of the respective years as well as a regression that helps visualize correlation and deviation from the median.
For example, when you select Yards as the data point, the graph will show the average yards per game of the top 10 NFL quarterback for each year from 2007- 2020 season.  As you can see yards per season have increased over the last decade and you can assume that if you are looking for a player to select in a draft, you would want to choose a player that has been able to keep up with the increase in trend and is performing near the current years average.  



```{r first, echo=FALSE, message=FALSE, warning=FALSE}
#Load the Library

library(shiny)
library(ggplot2)
library(plotly)
load(url("https://github.com/masaenz1/STA_6233_Spring2021/blob/main/Data_project_2/qbstatstotal.RData?raw=true"))

# Define UI for application that draws a histogram
ui <- fluidPage(
  
  # Application title
  titlePanel("QB STATS"),
  
  # Sidebar layout
  sidebarLayout(
    #Inputs: Select which inputs from the data we want to display
    sidebarPanel(
      #Select variable for y-axis
      selectInput(inputId = "y", 
                  label = "Y-axis:",
                  choices = c("YDS", "RTG", "TD", "QBR", "CMP"), 
                  selected = "YDS"),
      #Select X-axix variables
      selectInput(inputId = "x", 
                  label = "X-axis:",
                  choices = c("YEAR"), 
                  selected = "YEAR"),
      
      
      # Select Colors
      selectInput(inputId = "thecolors", 
                  label = "Choose Your Point Color",
                  choices = c("red", "green", "blue", "black"), 
                  selected = "black"),
    ),
    
    #Output: Type of plot
    mainPanel(
      plotlyOutput(outputId = "FreqTab") #Any name can go where "FreqTab" is, but be sure to keep it consistent with name in output$FreqTab in the server section
    )
  )
)

# Define server logic required to draw a histogram
server <- function(input, output) {
  
  output$FreqTab <- renderPlotly({
    # draw the histogram with the specified number of bins
    ggplot(qbstats, aes_string(x=input$x, y=input$y)) + geom_line( color="grey40") + geom_point(colour=input$thecolors, size=2, shape=23) + geom_smooth(method=lm, se=FALSE) + scale_x_continuous(breaks=c(2004,2007,2008,2009,2010,2011,2012,2013,2014,2015,2016,2017,2018,2019,2020))  
    
    
    #Notice the difference between the ggplots
  })
}
# Run the application 
shinyApp(ui = ui, server = server)
```
  
#### **Breakout of Statistics **  

Let’s take a closer look at the key stats and data points that you can use to identify the best player. Below are some additional visualizations as well as some explainations on each of the key stats and data points that aim to give you a better idea of what the stats arem current trends, and how they may be helpful to keep an eye on.   

```{r QB stat yards, echo=FALSE, warning=FALSE, message=FALSE}
#Load the Library

library(shiny)
library(ggplot2)
load(url("https://github.com/masaenz1/STA_6233_Spring2021/blob/main/Data_project_2/qbstatstotal.RData?raw=true"))

# Define UI for application that draws a histogram
ui <- fluidPage(
    # Application title
    titlePanel("Quarterback Stats- Yards per Season"),
    
    
    #Output: Type of plot
    mainPanel(
      plotOutput(outputId = "FreqTab") #Any name can go where "FreqTab" is, but be sure to keep it consistent with name in output$FreqTab in the server section
    )
  )

# Define server logic required to draw a histogram
server <- function(input, output) {
  
  output$FreqTab <- renderPlot({
    # draw the histogram with the specified number of bins
    ggplot(qbstats, aes(x=YEAR, y=YDS)) + geom_bar(stat = "identity", fill="aquamarine3") + geom_text(aes(label = YDS), vjust = -0.2, size=3) + ylim(0, 6000) +scale_x_continuous(breaks=c(2004,2007,2008,2009,2010,2011,2012,2013,2014,2015,2016,2017,2018,2019,2020))  #Notice the difference between the ggplots
  })
}
# Run the application 
shinyApp(ui = ui, server = server)

```


###### **Summary data points for YDS** 

```{r ui2, eval=TRUE, echo=FALSE, fig.align='default'}    
# Sidebar layout
summary(qbstats$YDS)

```

First, we look at yards per season.This statistic is one of the more basic stats but it can tell you a lot about a player.  As you can see from the bar chart below, yards per season has remained relatively consistent over the past decade and is currently trending upwards (e.g., the average yards per season in 2007 was 4,108 and in 2020 it was 4,460). You can also see that the mean for the past decade was 4,415 and can see that in 2020, the average of the top 10 quarerbacks were all putting up numbers above or just very close to that. 

This stat is really useful when looking for player to draft in fantasy leagues because the more yards your player puts up each game and each season, turns into more point that add up in your league.


```{r qbstat TD, echo=FALSE, warning=FALSE, message=FALSE}
#Load the Library

library(shiny)
library(ggplot2)
load(url("https://github.com/masaenz1/STA_6233_Spring2021/blob/main/Data_project_2/qbstatstotal.RData?raw=true"))

# Define UI for application that draws a histogram
ui <- fluidPage(
    # Application title
    titlePanel("Quarterback Stats- Touchdowns per Season"),
    
    
    #Output: Type of plot
    mainPanel(
      plotOutput(outputId = "FreqTab") #Any name can go where "FreqTab" is, but be sure to keep it consistent with name in output$FreqTab in the server section
    )
  )

# Define server logic required to draw a histogram
server <- function(input, output) {
  
  output$FreqTab <- renderPlot({
    # draw the histogram with the specified number of bins
    ggplot(qbstats, aes(x=YEAR, y=TD)) + geom_bar(stat = "identity", fill="azure4") + geom_text(aes(label = TD), vjust = -0.2, size=4) + ylim(0, 45) +scale_x_continuous(breaks=c(2004,2007,2008,2009,2010,2011,2012,2013,2014,2015,2016,2017,2018,2019,2020))  #Notice the difference between the ggplots
  })
}
# Run the application 
shinyApp(ui = ui, server = server)
```


###### **Summary data points for Touchdowns** 
```{r summary TD, eval=TRUE, echo=FALSE}    
# Sidebar layout
summary(qbstats$TD)
```
The next stat that we look at is the touchdowns per season data point.  This is another big stat especially when it comes to choosing player for fantasy leagues.  The more touchdowns your player can throw translates into more points that you get in your league.  It is the quickest way for your player to rack up points in each game.  You can see that this stat also shows an increasing trend (refer to interactive graph at top with correlation line) over the past decade.  The average touchdowns per season was around 30 touchdowns over the past dace but you can see that the top 10 players were averaging approximately 5 more in 2020 (at 35.2).  



```{r qbstat completion, echo=FALSE, warning=FALSE, message=FALSE}
#Load the Library

library(shiny)
library(ggplot2)
load(url("https://github.com/masaenz1/STA_6233_Spring2021/blob/main/Data_project_2/qbstatstotal.RData?raw=true"))

# Define UI for application that draws a histogram
ui <- fluidPage(
    # Application title
    titlePanel("Quarterback Stats- Completion Percentage per Season"),
    
    
    #Output: Type of plot
    mainPanel(
      plotOutput(outputId = "FreqTab") #Any name can go where "FreqTab" is, but be sure to keep it consistent with name in output$FreqTab in the server section
    )
  )

# Define server logic required to draw a histogram
server <- function(input, output) {
  
  output$FreqTab <- renderPlot({
    # draw the histogram with the specified number of bins
    ggplot(qbstats, aes(x=YEAR, y=CMP)) + geom_bar(stat = "identity", fill="deepskyblue4") + geom_text(aes(label = CMP), vjust = -0.2, size=3) + ylim(0, 75) +scale_x_continuous(breaks=c(2004,2007,2008,2009,2010,2011,2012,2013,2014,2015,2016,2017,2018,2019,2020))  #Notice the difference between the ggplots
  })
}
# Run the application 
shinyApp(ui = ui, server = server)
```


###### **Summary data points for Completion Percentages** 
```{r summary CMP, eval=TRUE, echo=FALSE}    
# Sidebar layout
summary(qbstats$CMP)
```
The average completion percentage for a quarterback is another key stat that show the accuracy of a quarterback.  The more accurate a quarterback can be a leading indicator of how well he will play over the season.   You can see that the completion percentage are relatively stable over the past decade but seeing a slight increasing trend.  The average over the past decade was jut under 65 percent but as of 2020, the top quarterbacks were putting up average percentages near 68 percent.  




```{r qbstat rating, echo=FALSE, warning=FALSE, message=FALSE}
#Load the Library

library(shiny)
library(ggplot2)
load(url("https://github.com/masaenz1/STA_6233_Spring2021/blob/main/Data_project_2/qbstatstotal.RData?raw=true"))

# Define UI for application that draws a histogram
ui <- fluidPage(
    # Application title
    titlePanel("Quarterback Stats- Passer Ratings"),
    
    
    #Output: Type of plot
    mainPanel(
      plotOutput(outputId = "FreqTab") #Any name can go where "FreqTab" is, but be sure to keep it consistent with name in output$FreqTab in the server section
    )
  )

# Define server logic required to draw a histogram
server <- function(input, output) {
  
  output$FreqTab <- renderPlot({
    # draw the histogram with the specified number of bins
    ggplot(qbstats, aes(x=YEAR, y=RTG)) + geom_bar(stat = "identity", fill="darkseagreen") + geom_text(aes(label = RTG), vjust = -0.2, size=3) + ylim(0, 110) +scale_x_continuous(breaks=c(2004,2007,2008,2009,2010,2011,2012,2013,2014,2015,2016,2017,2018,2019,2020))  #Notice the difference between the ggplots
  })
}
# Run the application 
shinyApp(ui = ui, server = server)
```


###### **Summary data points for QB Ratigns** 
```{r summary RTG, eval=TRUE, echo=FALSE}    
# Sidebar layout
summary(qbstats$RTG)
```
Another key stat and data point that people analyze over is the quarterback passer rating.  The passer rating is a key measure of the performance of a quarterback.  It is calculated using a players passing attempts, completions, yards, touchdowns, and interceptions with 0 being the worst and 158.3 being the best.  This stat is used by the NFL to determine its passing leaders for each season. 

As you can see, 2020 saw the most competition with the top player averaging a passer rating of 105.02, which is significantly better that the average over the past decade and a sign that quarterback are becoming increasingly better. 







```{r qbstat QBR, echo=FALSE, warning=FALSE, message=FALSE}
#Load the Library

library(shiny)
library(ggplot2)
load(url("https://github.com/masaenz1/STA_6233_Spring2021/blob/main/Data_project_2/qbstatstotal.RData?raw=true"))

# Define UI for application that draws a histogram
ui <- fluidPage(
    # Application title
    titlePanel("Quarterback Stats- QBR "),
    
    
    #Output: Type of plot
    mainPanel(
      plotOutput(outputId = "FreqTab") #Any name can go where "FreqTab" is, but be sure to keep it consistent with name in output$FreqTab in the server section
    )
  )

# Define server logic required to draw a histogram
server <- function(input, output) {
  
  output$FreqTab <- renderPlot({
    # draw the histogram with the specified number of bins
    ggplot(qbstats, aes(x=YEAR, y=QBR)) + geom_bar(stat = "identity", fill="coral3") + geom_text(aes(label = QBR), vjust = -0.2, size=3) + ylim(0, 75) +scale_x_continuous(breaks=c(2004,2007,2008,2009,2010,2011,2012,2013,2014,2015,2016,2017,2018,2019,2020))  #Notice the difference between the ggplots
  })
}
# Run the application 
shinyApp(ui = ui, server = server)
```


###### **Summary data points for QBR specific Ratigns** 
```{r summary QBR, eval=TRUE, echo=FALSE}    
# Sidebar layout
summary(qbstats$QBR)
```
The last stat or data point that we pull in is the “QBR” which is a relatively new stat that more and more people have been focusing on.  It is a stat that was created by ESPN and is similar to the passer rating stat but looks to go into more depth and detail to determine which quarterbacks really are the top player. The stat is a measure of performance but takes into account all potential factor that a quarterback may face such as yards, penalties, rushing yard, turnovers, and if they win the game.  

This stat can be used in conjunction with the other stat to help determine which player are the best all-around players.  As you can see the stat has been stable over the past decade with the max average coming in 2009 where the top quarterbacks record an average QBR of 73.66 while in 2020 the average the top player put up was right below at 72.77.  Either way, this stat is a good one to look at to determine all-around performance and if you are looking to draft a steady and reliable quarterback that will give you steady points throughout the year.



